---
title: "scRNASeq Lee - RENIA"
author: "Dana Krauss"
output:
  rmdformats::robobook:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: true
    highlight: tango
  html_document:
    toc: yes
    toc_depth: 4
    df_print: paged
    keep_md: yes
  html_notebook:
    theme: spacelab
    highlight: textmate
    toc: yes
    toc_depth: 4
    df_print: paged
    toc_float: yes
params: 
  prefix : "Lee_scRNASeq" 
  fdr_cutoff : 0.1
  fc_cutoff : 1
  wd: !r c("/srv/workspace/rprojects/USER/scRNASeq_Lee")
  script_number : !r c("04-")
---


***scRNASeq Analysis*** Scripts combined from Seurat workflow George/ICBI & <https://hbctraining.github.io/scRNA-seq_online/schedule/links-to-lessons.html>

***Dataset to analyse***

Analysis of Published scRNASeq Dataset <https://doi.org/10.1038/s41588-020-0636-z> Lineage-dependent gene expression programs influence the immune landscape of colorectal cancer

-   2 cohorts: KUL3 dataset , SMC dataset (10x Genomics)
-   91,103 unsorted single cells from 23 Korean & 6 Belgian


***
# Load libraries

<details>
  <summary>***Libraries & Params - Click to expand***</summary>
```{r setup}
knitr::opts_chunk$set(message = TRUE)
# load required libraries
library(knitr)
library("tidyverse")
library("conflicted")
library("BiocParallel")
library("ggplot2")
library("readr")
library("RColorBrewer")
library("dplyr")
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("count", "dplyr")
conflict_prefer("rowRanges", "MatrixGenerics")
library("ggpubr")
library("tibble")
conflict_prefer("paste", "base")
conflict_prefer("rename", "dplyr")
library("tidyr")
library("org.Hs.eg.db") #library("org.Mm.eg.db") 
conflict_prefer("desc", "dplyr")

library(Seurat)
library(Matrix)
library(RCurl)
library(scales)
library(cowplot)
library(SingleCellExperiment)
library(AnnotationHub)
library(ensembldb)
library(multtest)
library(glmGamPoi)
library(metap)
library(patchwork)

library(sctransform)
library(SingleR)
library(MAST)
library(DESeq2)
library(Nebulosa)
library(celldex)
library(ggrepel)
library("ComplexHeatmap")
library("circlize")


wd <- params$wd
results_dir = paste0(wd,"/results")
plots_dir = paste0(wd,"/plots")
data_dir =  paste0(wd,"/data")

### pull variables from params into environment
prefix <- params$prefix
conditions <- params$conditions
fdr_cutoff <- params$fdr_cutoff
fc_cutoff <- params$fc_cutoff
split_object <- params$split_integrate

# select which apply to data
color <- params$color
color_selection <- params$color_selection
n_cpus <- params$n_cpus
script_number <- params$script_number




save_plot <- function(filename, p, width=NULL, height=NULL) {
  if (!is.null(width) && !is.null(height)) {
    #ggsave(file.path(plots_dir, paste0("/",prefix,"/",filename, ".png")), plot = p, width = width, height = height)
   # ggsave(file.path(plots_dir, paste0("/",prefix,"/",filename, ".svg")), plot = p, width = width, height = height)
    ggsave(file.path(plots_dir, paste0("/",prefix,"/",filename, ".pdf")), plot = p, width = width, height = height)
  } else {
   # ggsave(file.path(plots_dir, paste0("/",prefix,"/",filename, ".png")), plot = p)
   # ggsave(file.path(plots_dir, paste0("/",prefix,"/",filename, ".svg")), plot = p)
    ggsave(file.path(plots_dir, paste0("/",prefix,"/",filename, ".pdf")), plot = p)
  }
}
save_RDS <- function(dir, object,filename){
  saveRDS(object, file.path(
    dir,paste0(prefix,"/",script_number, prefix ,"_",filename,".RDS")))
}


plotComposition <- function(seurat) {
table_samples_by_clusters <- seurat@meta.data %>%
  group_by(sample, cluster) %>%
  summarize(count = n()) %>%
  spread(cluster, count, fill = 0) %>%
  ungroup() %>%
  mutate(total_cell_count = rowSums(.[c(2:ncol(.))])) %>%
  dplyr::select(c('sample', 'total_cell_count', everything())) %>%
  arrange(factor(sample, levels = levels(seurat@meta.data$sample)))

knitr::kable(table_samples_by_clusters)

table_clusters_by_samples <- seurat@meta.data %>%
  group_by(cluster, sample) %>%
  summarize(count = n()) %>%
  spread(sample, count, fill = 0) %>%
  ungroup() %>%
  mutate(total_cell_count = rowSums(.[c(2:ncol(.))])) %>%
  select(c('cluster', 'total_cell_count', everything())) %>%
  arrange(factor(cluster, levels = levels(seurat@meta.data$cluster)))

knitr::kable(table_clusters_by_samples)

temp_labels <- seurat@meta.data %>%
  group_by(sample) %>%
  tally()


p1 <- table_samples_by_clusters %>%
  select(-c("total_cell_count")) %>%
  reshape2::melt(id.vars = "sample") %>%
  mutate(sample = factor(sample, levels = levels(as.factor(seurat@meta.data$sample)))) %>%
  ggplot(aes(sample, value)) +
  geom_bar(aes(fill = variable), position = "stack", stat = "identity") +
  geom_text(
    data = temp_labels,
    aes(x = sample, y = Inf, label = paste0('n = ', format(n, big.mark = ',', trim = TRUE)), angle = 45, hjust= 0, vjust=-1),
    color = 'black', size = 2.8
  ) +
  scale_fill_manual(name = "Cluster"
                    , values = custom_colors$discrete
                    ) +
  scale_y_continuous(name = "Number of cells", labels = scales::comma, expand = c(0.01, 0)) +
  coord_cartesian(clip = 'off') +
  theme_bw() +
  theme(
    legend.position = 'left',
    plot.title = element_text(hjust = 0.5),
    text = element_text(size = 16),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    plot.margin = margin(t = 20, r = 0, b = 0, l = 0, unit = 'pt')
  )

temp_labels <- seurat@meta.data %>%
  group_by(cluster) %>%
  tally() %>%
  dplyr::rename('cluster' = cluster)

p2 <- table_clusters_by_samples %>%
  select(-c('total_cell_count')) %>%
  reshape2::melt(id.vars = 'cluster') %>%
  mutate(cluster = factor(cluster, levels = levels(seurat@meta.data$cluster))) %>%
  ggplot(aes(cluster, value)) +
  geom_bar(aes(fill = variable), position = 'stack', stat = 'identity') +
  geom_text(
    data = temp_labels,
    aes(x = cluster, y = Inf, label = paste0('n = ', format(n, big.mark = ',', trim = TRUE)), vjust = -1),
    color = 'black', size = 2.8
  ) +
  scale_fill_manual(name = 'sample'
                    ,values =  custom_colors$discrete[2:3]
                    ) +
  scale_y_continuous(labels = scales::comma, expand = c(0.01, 0)) +
  coord_cartesian(clip = 'off') +
  theme_bw() +
  theme(
    legend.position = 'right',
    plot.title = element_text(hjust = 0.5),
    text = element_text(size = 16),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_blank(),
    plot.margin = margin(t = 20, r = 0, b = 0, l = 10, unit = 'pt')
  )
p1 + p2 +
    ggtitle("Composition SampleClusters by Number") +
    theme(plot.title = element_text(color="black", size=14, vjust = 5)) +
    plot_layout(ncol = 2, widths = c(
        seurat@meta.data$sample %>% unique() %>% length(),
        seurat@meta.data$cluster %>% unique() %>% length()
    ))

p_number <- p1 + p2 +
    ggtitle("Composition SampleClusters by Number") +
    theme(plot.title = element_text(color="black", size=14, vjust = 5)) +
    plot_layout(ncol = 2, widths = c(
        seurat@meta.data$sample %>% unique() %>% length(),
        seurat@meta.data$cluster %>% unique() %>% length()
    ))
save_plot(paste0(script_number,prefix,"_",title,"_composition_samples_cluster_by_number"), p_number, width = 18, height = 8)


temp_labels <- seurat@meta.data %>%
  group_by(sample) %>%
  tally()

p3 <- table_samples_by_clusters %>%
  select(-c('total_cell_count')) %>%
  reshape2::melt(id.vars = 'sample') %>%
  mutate(sample = factor(sample, levels = levels((as.factor(seurat@meta.data$sample))))) %>%
  ggplot(aes(sample, value)) +
  geom_bar(aes(fill = variable), position = 'fill', stat = 'identity') +
  geom_text(
    data = temp_labels,
    aes(x = sample, y = Inf, label = paste0('n = ', format(n, big.mark = ',', trim = TRUE)), angle = 45, hjust= 0, vjust=-1),
    color = 'black', size = 2.8
  ) +
  scale_fill_manual(name = 'Cluster', values = custom_colors$discrete) +
  scale_y_continuous(name = 'Percentage [%]', labels = scales::percent_format(), expand = c(0.01,0)) +
  coord_cartesian(clip = 'off') +
  theme_bw() +
  theme(
    legend.position = 'left',
    plot.title = element_text(hjust = 0.5),
    text = element_text(size = 16),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    plot.margin = margin(t = 20, r = 0, b = 0, l = 0, unit = 'pt')
  )

temp_labels <- seurat@meta.data %>%
  group_by(cluster) %>%
  tally() %>%
  dplyr::rename('cluster' = cluster)
p4 <- table_clusters_by_samples %>%
  select(-c('total_cell_count')) %>%
  reshape2::melt(id.vars = 'cluster') %>%
  mutate(cluster = factor(cluster, levels = levels(seurat@meta.data$cluster))) %>%
  ggplot(aes(cluster, value)) +
  geom_bar(aes(fill = variable), position = 'fill', stat = 'identity') +
  geom_text(
    data = temp_labels, aes(x = cluster, y = Inf, label = paste0('n = ', format(n, big.mark = ',', trim = TRUE)), vjust = -1),
    color = 'black', size = 2.8
  ) +
  scale_fill_manual(name = 'sample', values = custom_colors$discrete[2:3]) +
  scale_y_continuous(name = 'Percentage [%]', labels = scales::percent_format(), expand = c(0.01,0)) +
  coord_cartesian(clip = 'off') +
  theme_bw() +
  theme(
    legend.position = 'right',
    plot.title = element_text(hjust = 0.5),
    text = element_text(size = 16),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_blank(),
    plot.margin = margin(t = 20, r = 0, b = 0, l = 10, unit = 'pt')
  )

p3 + p4 +
    ggtitle("Composition SampleClusters by Percentage ") +
    theme(plot.title = element_text(color="black", size=14, vjust = 5)) +
    plot_layout(ncol = 2, widths = c(
        seurat@meta.data$sample %>% unique() %>% length(),
        seurat@meta.data$cluster %>% unique() %>% length()
    ))

p_percent <- p3 + p4 +
    ggtitle("Composition SampleClusters by Percentage") +
    theme(plot.title = element_text(color="black", size=14, vjust = 5)) +
    plot_layout(ncol = 2, widths = c(
        seurat@meta.data$sample %>% unique() %>% length(),
        seurat@meta.data$cluster %>% unique() %>% length()
    ))

save_plot(paste0(script_number,prefix,"_",title,"_composition_samples_topLayer_by_percent"), p_percent, width = 18, height = 8)
return(list(p_number, p_percent))
}
```
 </details> 

Some custom colors
```{r}
custom_colors <- list()

colors_dutch <- c(
  '#FFC312','#C4E538','#12CBC4','#FDA7DF','#ED4C67',
  '#F79F1F','#A3CB38','#1289A7','#D980FA','#B53471',
  '#EE5A24','#009432','#0652DD','#9980FA','#833471',
  '#EA2027','#006266','#1B1464','#5758BB','#6F1E51'
)

colors_spanish <- c(
  '#40407a','#706fd3','#f7f1e3','#34ace0','#33d9b2',
  '#2c2c54','#474787','#aaa69d','#227093','#218c74',
  '#ff5252','#ff793f','#d1ccc0','#ffb142','#ffda79',
  '#b33939','#cd6133','#84817a','#cc8e35','#ccae62'
)

custom_colors$discrete <- c(colors_dutch, colors_spanish)

custom_colors$cell_cycle <- setNames(
  c('#45aaf2', '#f1c40f', '#e74c3c', '#7f8c8d'),
  c('G1',      'S',       'G2M',     '-')
)
```

# Read in data from before
```{r}
# read back in- seurat object normalised
seurat_obj_normalized <- readRDS(file.path(data_dir, paste0("03-", prefix, "_seurat_topLayer.RDS")))

seurat_obj_normalized <- NormalizeData(object = seurat_obj_normalized)
seurat_obj_normalized <- FindVariableFeatures(object = seurat_obj_normalized)
seurat_obj_normalized <- ScaleData(object = seurat_obj_normalized)

DefaultAssay(seurat_obj_normalized) <- "RNA"
seurat_obj_normalized <- NormalizeData(object = seurat_obj_normalized)
seurat_obj_normalized <- FindVariableFeatures(object = seurat_obj_normalized)
seurat_obj_normalized <- ScaleData(object = seurat_obj_normalized)

unique(Idents(seurat_obj_normalized))

#Idents(seurat_obj_normalized)
prop.table(table(Idents(seurat_obj_normalized), seurat_obj_normalized$KRAS), margin = 2)
```

## EGFR positive cells

```{r EGFR feature plots RNA}
p <- FeaturePlot(seurat_obj_normalized, features = "EGFR", 
                 max.cutoff = 1, 
                 pt.size = 0.01,
                 reduction = "umap",
    cols = c("lightgrey", "#04103C"),
    keep.scale = "feature",
    label = T,
    raster=F)
p
save_plot(paste0(script_number,prefix,"_umap_EGFR_all"), p, height = 10,width = 12)

p <- FeaturePlot(seurat_obj_normalized, features = "EGFR", 
                 max.cutoff = 1, 
                 pt.size = 0.01,
                 reduction = "tsne",
    cols = c("lightgrey", "#04103C"),
    keep.scale = "feature",
    label = T,
    raster=F)
p
save_plot(paste0(script_number,prefix,"_tsne_EGFR_all"), p, height = 10,width = 12)



#### Plot tsne & umap plots
Idents(seurat_obj_normalized) <- seurat_obj_normalized$topLayer
p1 <- DimPlot(seurat_obj_normalized, reduction = "tsne") + 
   ggtitle('Cell types top layer')
Idents(seurat_obj_normalized) <- seurat_obj_normalized$Cell_subtype
p2 <- DimPlot(seurat_obj_normalized, reduction = "tsne") + 
   ggtitle('Cell types subtype')
p1 | p2
p <- p1 + p2
save_plot(paste0(script_number,prefix,"_tsne"), p, height = 8,width = 20)

Idents(seurat_obj_normalized) <- seurat_obj_normalized$topLayer
p1 <- DimPlot(seurat_obj_normalized, reduction = "umap") + 
   ggtitle('Cell types top layer')
Idents(seurat_obj_normalized) <- seurat_obj_normalized$Cell_subtype
p2 <- DimPlot(seurat_obj_normalized, reduction = "umap") + 
   ggtitle('Cell types subtype')
p1 | p2
p <- p1 + p2
save_plot(paste0(script_number,prefix,"_umap"), p, height = 8,width = 20)




#violin EGFR RNA total
Idents(seurat_obj_normalized) <- seurat_obj_normalized$topLayer
library(scCustomize)
p <- VlnPlot_scCustom(seurat_object = seurat_obj_normalized,
                      features = "EGFR", 
                      raster = FALSE)
p  
save_plot(paste0(script_number,prefix,"_violin_EGFR_Expression"), p, height = 10,width = 16)
p <- VlnPlot(seurat_obj_normalized,  features = "EGFR" )
p  
save_plot(paste0(script_number,prefix,"_violin_EGFR_Expression_2"), p, height = 10,width = 16)

# vln_df = data.frame(EGFR = seurat_obj_normalized[["RNA"]]@data["EGFR",], cluster = seurat_obj_normalized$topLayer)
# ggplot(vln_df, aes(x = cluster, y = EGFR)) + geom_violin(aes(fill = cluster), trim=TRUE, scale = "width")





#Subset Myeloid 
seurat_myeloid <- subset(x = seurat_obj_normalized, idents = "Myeloid")
p <- FeaturePlot(seurat_myeloid, features = "EGFR", 
                 max.cutoff = 1, 
                 pt.size = 0.01,
                 reduction = "umap",
                 #split.by = "Health",
    #cols = c("grey", "midnightblue"),
    cols = c("lightgrey", "#04103C"),
    keep.scale = "feature",
    #blend = F,
    #cols = palette_blues,
    label = T,
    raster=F)
p
save_plot(paste0(script_number,prefix,"_umap_EGFR_Myeloid"), p, height = 10,width = 12)





```







## PROGENy
```{r PROGENy sc}
library(progeny)
library(decoupleR)
library(pheatmap)
net <- get_progeny(organism = 'human', top = 100)
net

# Extract the normalized log-transformed counts
data <- seurat_obj_normalized
mat <- as.matrix(data@assays$RNA@data)



##### WATCH OUT, this part runs very long!!! I save the result separately, you can laod it again a bin down 


# Run wmean
acts <- run_wmean(mat=mat, net=net, .source='source', .target='target',
                  .mor='weight', times = 100, minsize = 5)
acts

# Extract norm_wmean and store it in pathwayswmean in data
data[['pathwayswmean']] <- acts %>%
  dplyr::filter(statistic == 'norm_wmean') %>%
  pivot_wider(id_cols = 'source', names_from = 'condition',
              values_from = 'score') %>%
  column_to_rownames('source') %>%
  Seurat::CreateAssayObject(.)

# Change assay
DefaultAssay(object = data) <- "pathwayswmean"

# Scale the data
data <- ScaleData(data)
data@assays$pathwayswmean@data <- data@assays$pathwayswmean@scale.data




###### read in data separately

data <- readRDS(file.path(data_dir, paste0("04-", prefix, "seurat_egfr_info.RDS")))




#"lightgrey", "#04103C"
p1 <- DimPlot(data, reduction = "umap", label = TRUE, pt.size = 0.5, group.by = "topLayer") + 
  NoLegend() + ggtitle('Cell types')
p2 <- (FeaturePlot(data, features = c("EGFR")) & 
  scale_colour_gradient2(low = 'black', mid = 'lightgrey', high = 'red')) +
  ggtitle('EGFR activity')
p1 | p2
p <- p1 + p2
save_plot(paste0(script_number,prefix,"_EGFR_Pathway_Progeny"), p, height = 10,width = 16)


#Subset myeloid cells
Idents(data) <- data$Cell_subtype

seurat_myeloid <- subset(x = data, 
                         idents = c("SPP1+","Pro-inflammatory","Proliferating","cDC"))
Idents(seurat_myeloid) <- seurat_myeloid$topLayer
seurat_myeloid <- subset(x = seurat_myeloid, idents = c("Myeloid"))

p1 <- DimPlot(seurat_myeloid, reduction = "umap", label = TRUE, pt.size = 0.5, group.by = "topLayer") + 
  NoLegend() + ggtitle('Cell types')
p2 <- DimPlot(seurat_myeloid, reduction = "umap", label = TRUE, pt.size = 0.5, group.by = "Cell_subtype") + 
  NoLegend() + ggtitle('Cell types')
p3 <- (FeaturePlot(seurat_myeloid,  features = c("EGFR"), max.cutoff = 2,  pt.size = 0.5, reduction = "umap", label = T, raster=F) & scale_colour_gradient2(low = '#04103C', mid = 'lightgrey', high = '#e84445')) +
    ggtitle('EGFR activity')
p1 | p2 | p3


p <- p1 | p2 | p3
p
save_plot(paste0(script_number,prefix,"_EGFR_Pathway_Progeny_Myeloid"), p, height = 10,width = 18)


#Plot violin Progeny total
library(scCustomize)

p <- VlnPlot_scCustom(seurat_object = data, 
                      features = "EGFR", 
                      raster = FALSE,
                 pt.size = 0.01)
save_plot(paste0(script_number,prefix,"_violin_EGFR_Progeny_Myeloid"), p, height = 10,width = 16)

# Extract activities from object as a long dataframe
df <- t(as.matrix(data@assays$pathwayswmean@data)) %>%
  as.data.frame()
hist(df$EGFR, breaks = 1000)

# Extract activities from object as a long dataframe for myeloid cells
df <- t(as.matrix(seurat_myeloid@assays$pathwayswmean@data)) %>%
  as.data.frame()
hist(df$EGFR, breaks = 1000)

ordered_egfr = df$EGFR[order(df$EGFR,decreasing = F)]
plot(x = 1:length(ordered_egfr), y = ordered_egfr)

length(df$EGFR)

#table(df$EGFR)

length(which(df$EGFR >= 0))
df_ordered <- df %>% arrange(EGFR)
  
quantile(x = df_ordered$EGFR, type = 4)



egfr.cluster <- df %>%
  add_column(EGFR_positive = df$EGFR >= 0.77) %>%
  add_column(EGFR_negative = df$EGFR <= -0.52) %>%
  mutate(EGFR_pathway = case_when(
    EGFR_positive == TRUE ~ "up",
    EGFR_negative == TRUE ~ "down",
    EGFR_positive == FALSE ~ "else"))
all(rownames(egfr.cluster) == rownames(seurat_myeloid@meta.data))

seurat_myeloid$EGFR_pathway_activity <- egfr.cluster$EGFR_pathway
saveRDS(seurat_myeloid, file.path(data_dir, paste0(script_number,prefix, "seurat_myeloid_egfr_pathway.RDS")))

############## plot again after classification
DimPlot(seurat_myeloid, 
        pt.size = 0.5,
        reduction = "umap", 
        group.by = "EGFR_pathway_activity",
        label = F,
        cols = c("#d9d9d9","lightgrey","#ff0000" )
        ) &
    ggtitle('EGFR activity')




p1 <- DimPlot(seurat_myeloid, 
              reduction = "umap", 
              label = TRUE, pt.size = 0.5, 
              group.by = "topLayer") + 
  NoLegend() + ggtitle('Cell types')
p2 <- DimPlot(seurat_myeloid, 
              reduction = "umap", 
              label = TRUE, 
              pt.size = 0.5, 
              group.by = "Cell_subtype") + 
  NoLegend() + ggtitle('Cell types')
p3 <- DimPlot(seurat_myeloid, 
              pt.size = 0.5,
              reduction = "umap", 
              group.by = "EGFR_pathway_activity",
              label = F,
              cols = c("#d9d9d9","lightgrey","#ff0000")) &
    ggtitle('EGFR activity')
p1 | p2 | p3

p <- p1 | p2 | p3
save_plot(paste0(script_number,prefix,"_EGFR_Pathway_Progeny_Myeloid_classification"), p, height = 10,width = 18)

prop.table(table(seurat_myeloid$EGFR_pathway_activity,seurat_myeloid$Cell_subtype ), margin = 2)









# Extract activities from object as a long dataframe
df <- t(as.matrix(data@assays$pathwayswmean@data)) %>%
  as.data.frame() %>%
  mutate(cluster = Idents(data)) %>%
  pivot_longer(cols = -cluster, names_to = "source", values_to = "score") %>%
  group_by(cluster, source) %>%
  summarise(mean = mean(score))

# Transform to wide matrix
top_acts_mat <- df %>%
  pivot_wider(id_cols = 'cluster', names_from = 'source',
              values_from = 'mean') %>%
  column_to_rownames('cluster') %>%
  as.matrix()

# Choose color palette
palette_length = 100
my_color = colorRampPalette(c("Darkblue", "white","red"))(palette_length)

my_breaks <- c(seq(-2, 0, length.out=ceiling(palette_length/2) + 1),
               seq(0.05, 2, length.out=floor(palette_length/2)))

# Plot
pheatmap::pheatmap(top_acts_mat, border_color = NA, color=my_color, breaks = my_breaks)





saveRDS(data, file.path(data_dir, paste0(script_number,prefix, "seurat_egfr_info.RDS")))



```



## Subcluster Myeloid

```{r Subcluster Myeloid}
# DefaultAssay(seurat_myeloid) <- "SCT"
# seurat_myeloid <- RunPCA(object = seurat_myeloid, 
#                    npcs = 30, 
#                    features = VariableFeatures(object = seurat_myeloid))
# ElbowPlot(seurat_myeloid, ndims = 30)
# seurat_myeloid <- FindNeighbors(object = seurat_myeloid, 
#                                 dims = 1:4) #22? #adjust 1:40
# seurat_myeloid <- FindClusters(object = seurat_myeloid, 
#                                resolution = c(0.4, 0.5, 0.6, 0.7,0.8)) #c(0.4, 0.6, 0.8, 1.0, 1.4) #0.55
# 
# unique(seurat_myeloid$SCT_snn_res.0.4)
# Idents(seurat_myeloid) <- seurat_myeloid$SCT_snn_res.0.4
# 
# seurat_myeloid <- RunUMAP(object = seurat_myeloid,
#                     dims = 1:6,#adjust 11? # 25?
#                     reduction = "pca",
#                     n.neighbors = 26,
#                     n.epochs = 400)
# 
# seurat_myeloid <- RunTSNE(object = seurat_myeloid, 
#                     dims = 1:6, #adjust 11? 25?
#                     perplexity = 30, #default = 30
#                     max_iter = 1000, # max iteration, default = 1000
#                     eta = 200, #learning rate, default = 200
#                     verbose = TRUE) 
# saveRDS(seurat_myeloid, file.path(data_dir, paste0(script_number,prefix, "seurat_myeloid.RDS")) )

DimPlot(seurat_myeloid, group.by = "Cell_subtype")
DimPlot(seurat_myeloid, group.by = "Cell_subtype", reduction = "tsne")
FeaturePlot(seurat_myeloid, features = c("CD163", "SEPP1", "APOE", "MAF", "SPP1", "MKI67", "STMN1", "IL1A", "IL1B"), label = T)




p1 <- DimPlot(seurat_myeloid, 
              reduction = "umap", 
              label = TRUE, pt.size = 0.5, 
              group.by = "topLayer") + 
  NoLegend() + ggtitle('Cell types')
p2 <- DimPlot(seurat_myeloid, 
              reduction = "umap", 
              label = TRUE, 
              pt.size = 0.5, 
              group.by = "Cell_subtype") + 
  NoLegend() + ggtitle('Cell types')
p3 <- DimPlot(seurat_myeloid, 
              pt.size = 0.5,
              reduction = "umap", 
              group.by = "EGFR_pathway_activity",
              label = F,
              cols = c("#d9d9d9","lightgrey","#ff0000")) &
    ggtitle('EGFR activity')
p <- p1 | p2 | p3
save_plot(paste0(script_number,prefix,"_EGFR_Pathway_Progeny_Myeloid_classification_UMAP"), p, height = 10,width = 18)



p1 <- DimPlot(seurat_myeloid, 
              reduction = "tsne", 
              label = TRUE, pt.size = 0.5, 
              group.by = "topLayer") + 
  NoLegend() + ggtitle('Cell types')
p2 <- DimPlot(seurat_myeloid, 
              reduction = "tsne", 
              label = TRUE, 
              pt.size = 0.5, 
              group.by = "Cell_subtype") + 
  NoLegend() + ggtitle('Cell types')
p3 <- DimPlot(seurat_myeloid, 
              pt.size = 0.5,
              reduction = "tsne", 
              group.by = "EGFR_pathway_activity",
              label = F,
              cols = c("#d9d9d9","lightgrey","#ff0000")) &
    ggtitle('EGFR activity')
p <- p1 | p2 | p3
save_plot(paste0(script_number,prefix,"_EGFR_Pathway_Progeny_Myeloid_classification_tsne"), p, height = 10,width = 18)


prop.table(table(seurat_myeloid$EGFR_pathway_activity,seurat_myeloid$Cell_subtype ), margin = 2)
```



### DEGs Myeloid

```{r DE Myeloid Analysis Pathway up or down}
seurat_DE <- seurat_myeloid
Idents(seurat_DE) <- seurat_DE$EGFR_pathway_activity

seurat_DE_HvL <- subset(seurat_DE, idents = c("up", "down"))

#Comparison for genes happens here (EGFRlovshi Activity)
DefaultAssay(seurat_DE_HvL) <- "RNA"
diff_markers_all <- FindMarkers(object = seurat_DE_HvL,
                                     ident.1 = "down",
                                     ident.2 = "up",
                                     logfc.threshold = 0.01,
                                     only.pos = FALSE) 
diff_markers_all <- diff_markers_all %>%
  rownames_to_column("GeneName")

library(writexl)
write_xlsx(diff_markers_all, file.path(results_dir, paste0(script_number,prefix, "_down_vs_up_progeny_EGFR_deg_genes_FindMarkers_wilcox" , ".xlsx")))





Idents(seurat_DE_HvL) <- seurat_DE_HvL$Cell_subtype
diff_markers_Proinflammatory <- FindMarkers(object = seurat_DE_HvL,
                                     ident.1 = "down",
                                     ident.2 = "up",
                                     group.by = "EGFR_pathway_activity",
                                     logfc.threshold = 0.01,
                                     only.pos = FALSE,
                                     subset.ident = "Pro-inflammatory")
diff_markers_Proinflammatory <- diff_markers_Proinflammatory %>%
  rownames_to_column("GeneName")
write_xlsx(diff_markers_Proinflammatory, file.path(results_dir, paste0(script_number,prefix, "_down_vs_up_deg_Proinflammatory_wilcox" , ".xlsx")))

diff_markers_SPP1 <- FindMarkers(object = seurat_DE_HvL,
                                     ident.1 = "down",
                                     ident.2 = "up",
                                     group.by = "EGFR_pathway_activity",
                                     logfc.threshold = 0.01,
                                     only.pos = FALSE,
                                     subset.ident = "SPP1+")
diff_markers_SPP1 <- diff_markers_SPP1 %>%
  rownames_to_column("GeneName")
write_xlsx(diff_markers_SPP1, file.path(results_dir, paste0(script_number,prefix, "_down_vs_up_deg_SPP1_wilcox" , ".xlsx")))



diff_markers_Proliferating <- FindMarkers(object = seurat_DE_HvL,
                                     ident.1 = "down",
                                     ident.2 = "up",
                                     group.by = "EGFR_pathway_activity",
                                     logfc.threshold = 0.01,
                                     only.pos = FALSE,
                                     subset.ident = "Proliferating")
diff_markers_Proliferating <- diff_markers_Proliferating %>%
  rownames_to_column("GeneName")
write_xlsx(diff_markers_Proliferating, file.path(results_dir, paste0(script_number,prefix, "_down_vs_up_deg_Proliferating_wilcox" , ".xlsx")))



diff_markers_collected <- list(
         down_vs_up_all_egfr_pathway = diff_markers_all,
         down_vs_up_Proinflammatory_egfr_pathway = diff_markers_Proinflammatory,
         down_vs_up_SPP1_egfr_pathway = diff_markers_SPP1,
         down_vs_up_Proliferating_egfr_pathway = diff_markers_Proliferating)

write_xlsx(diff_markers_collected, file.path(results_dir, paste0(script_number,prefix, "_collected_deg_down_vs_up_progeny_egfr" , ".xlsx")))



###Create Volcano plot for myeloid (EGFRlo vs hi activity)- using a treshold

# add a column of NAs
diff_markers_SPP1$diffexpressed <- "NO"
# if log2Foldchange > 0.5 and pvalue < 0.05, set as "UP" 
diff_markers_SPP1$diffexpressed[diff_markers_SPP1$avg_log2FC > 0.5 & diff_markers_SPP1$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.5 and pvalue < 0.05, set as "DOWN"
diff_markers_SPP1$diffexpressed[diff_markers_SPP1$avg_log2FC < -0.5 & diff_markers_SPP1$p_val_adj < 0.05] <- "DOWN"



# Now write down the name of genes beside the points...
# Create a new column "delabel" to de, that will contain the name of genes differentially expressed (adda NA in case they are not)
diff_markers_SPP1$delabel <- NA
diff_markers_SPP1$delabel[diff_markers_SPP1$diffexpressed !="NO" ] <- diff_markers_SPP1$GeneName[diff_markers_SPP1$diffexpressed != "NO"]


#Draw Volcano
q<-ggplot(data=diff_markers_SPP1, aes(x=avg_log2FC, y=-log10(p_val_adj), col=diffexpressed, label= delabel)) + 
    geom_point() + 
    theme_minimal()+
  geom_text_repel(max.overlaps = Inf)
q

# Add lines as before...
q2 <- q + geom_vline(xintercept=c(--0.5, -0.5), col="red") + xlim(-2, 2)+ 
        geom_hline(yintercept=-log10(0.05), col="red") +  ylim(0, 75) 
q2

## Change point color 

# 2. to automate a bit: ceate a named vector: the values are the colors to be used, the names are the categories they will be assigned to:
mycolors <- c("blue", "red", "black")
names(mycolors) <- c("DOWN", "UP", "NO")
q3 <- q2 + scale_colour_manual(values = mycolors)

q3

save_plot(paste0(script_number,prefix,"_Volcano_SPP1_LOvsHi_limit_y"), q3, height = 10,width = 10)
```


#Create volcano with selected genes
```{r New myeloid volcanos selection of genes}

###Create Volcano plot for myeloid (EGFRlo vs hi activity) using selected genes
# add a column of NAs
diff_markers_SPP1$diffexpressed <- "NO"
# if log2Foldchange > 0.5 and pvalue < 0.05, set as "UP" 
diff_markers_SPP1$diffexpressed[diff_markers_SPP1$avg_log2FC > 0 & diff_markers_SPP1$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.5 and pvalue < 0.05, set as "DOWN"
diff_markers_SPP1$diffexpressed[diff_markers_SPP1$avg_log2FC < -0 & diff_markers_SPP1$p_val_adj < 0.05] <- "DOWN"



# Create a new column "delabel" to de, that will contain the name of genes differentially expressed (adda NA in case they are not)
diff_markers_SPP1$delabel <- NA
diff_markers_SPP1$delabel[diff_markers_SPP1$diffexpressed !="NO" ] <- diff_markers_SPP1$GeneName[diff_markers_SPP1$diffexpressed != "NO"]


diff_markers_SPP1$genelabels <- ifelse(diff_markers_SPP1$GeneName == "IL1A" 
                               |diff_markers_SPP1$GeneName == "IL1B"
                              |diff_markers_SPP1$GeneName == "IL1RN"
                              |diff_markers_SPP1$GeneName == "CXCL8"
                              |diff_markers_SPP1$GeneName == "DUSP4"
                              |diff_markers_SPP1$GeneName == "CCL3"
                              |diff_markers_SPP1$GeneName == "CXCL2"
                              |diff_markers_SPP1$GeneName == "CXCL3"
                              |diff_markers_SPP1$GeneName == "CCL20"
                              |diff_markers_SPP1$GeneName == "NFKBIA"	
                              |diff_markers_SPP1$GeneName == "NFKB1"
                              |diff_markers_SPP1$GeneName == "NLRP3"
                              |diff_markers_SPP1$GeneName == "CCL2"
                              |diff_markers_SPP1$GeneName == "MAP2K3"
                              |diff_markers_SPP1$GeneName == "CDKN1A"	
                              |diff_markers_SPP1$GeneName == "EREG"
                              |diff_markers_SPP1$GeneName == "AREG"
                               |diff_markers_SPP1$GeneName == "PLEK"
                               |diff_markers_SPP1$GeneName == "CCL4"
                              |diff_markers_SPP1$GeneName == "TNF"
                              |diff_markers_SPP1$GeneName == "G0S2"
                               |diff_markers_SPP1$GeneName == "APOE"
                               |diff_markers_SPP1$GeneName == "APOC1"
                              |diff_markers_SPP1$GeneName == "C1QA"
                              |diff_markers_SPP1$GeneName == "C1QB"
                              |diff_markers_SPP1$GeneName == "NUPR1"
                              |diff_markers_SPP1$GeneName == "CXCL9"
                               |diff_markers_SPP1$GeneName == "CCL18",
                              TRUE, FALSE)  
  
#Draw Volcano
q<-ggplot(data=diff_markers_SPP1, aes(x=avg_log2FC, y=-log10(p_val_adj), col=diffexpressed # label= delabel)) +
                                      ))+ 
      geom_point() + 
    theme_minimal()+
  geom_text_repel(
     aes(avg_log2FC, -log10(p_val_adj)),
     label = ifelse(diff_markers_SPP1$genelabels, diff_markers_SPP1$GeneName, ""),
     box.padding = unit(0.45, "lines"),
     hjust = 1, max.overlaps = Inf) 


q

# Add lines as before...
q2 <- q + geom_vline(xintercept=c(--0, -0), col="black") + xlim(-2, 2)+ 
        geom_hline(yintercept=-log10(0.05), col="black") +  ylim(0, 75) 
q2

# 2. to automate a bit: ceate a named vector: the values are the colors to be used, the names are the categories they will be assigned to:
mycolors <- c("blue", "red", "black")
names(mycolors) <- c("DOWN", "UP", "NO")
q3 <- q2 + scale_colour_manual(values = mycolors)

q3

save_plot(paste0(script_number,prefix,"_Volcano_SPP1_LOvsHi_selec_limit_y"), q3, height = 8,width = 8)



#Pro-inflammatory
###Create Volcano plot for myeloid (EGFRlo vs hi activity)
# add a column of NAs
diff_markers_Proinflammatory$diffexpressed <- "NO"
# if log2Foldchange > 0.5 and pvalue < 0.05, set as "UP" 
diff_markers_Proinflammatory$diffexpressed[diff_markers_Proinflammatory$avg_log2FC > 0 & 
                               diff_markers_Proinflammatory$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.5 and pvalue < 0.05, set as "DOWN"
diff_markers_Proinflammatory$diffexpressed[diff_markers_Proinflammatory$avg_log2FC < -0 & diff_markers_Proinflammatory$p_val_adj < 0.05] <- "DOWN"



# Create a new column "delabel" to de, that will contain the name of genes differentially expressed (adda NA in case they are not)
diff_markers_Proinflammatory$delabel <- NA
diff_markers_Proinflammatory$delabel[diff_markers_Proinflammatory$diffexpressed !="NO" ] <- diff_markers_Proinflammatory$GeneName[diff_markers_Proinflammatory$diffexpressed != "NO"]


diff_markers_Proinflammatory$genelabels <- ifelse(diff_markers_Proinflammatory$GeneName == "IL1A" 
                               |diff_markers_Proinflammatory$GeneName == "IL1B"
                              |diff_markers_Proinflammatory$GeneName == "IL1RN"
                              |diff_markers_Proinflammatory$GeneName == "CXCL8"
                              |diff_markers_Proinflammatory$GeneName == "DUSP4"
                              |diff_markers_Proinflammatory$GeneName == "CCL3"
                              |diff_markers_Proinflammatory$GeneName == "CXCL2"
                              |diff_markers_Proinflammatory$GeneName == "CXCL3"
                              |diff_markers_Proinflammatory$GeneName == "CCL20"
                              |diff_markers_Proinflammatory$GeneName == "NFKBIA"	
                              |diff_markers_Proinflammatory$GeneName == "NFKB1"
                              |diff_markers_Proinflammatory$GeneName == "NLRP3"
                              |diff_markers_Proinflammatory$GeneName == "CCL2"
                              |diff_markers_Proinflammatory$GeneName == "MAP2K3"
                              |diff_markers_Proinflammatory$GeneName == "CDKN1A"	
                              |diff_markers_Proinflammatory$GeneName == "EREG"
                              |diff_markers_Proinflammatory$GeneName == "AREG"
                               |diff_markers_Proinflammatory$GeneName == "PLEK"
                               |diff_markers_Proinflammatory$GeneName == "CCL4"
                              |diff_markers_Proinflammatory$GeneName == "TNF"
                              |diff_markers_Proinflammatory$GeneName == "VCAN"
                               |diff_markers_Proinflammatory$GeneName == "THBS1"
                              |diff_markers_Proinflammatory$GeneName == "S100A6"
                               |diff_markers_Proinflammatory$GeneName == "S100A8"
                              |diff_markers_Proinflammatory$GeneName == "S100A9"
                               |diff_markers_Proinflammatory$GeneName == "ZEB2"
                               |diff_markers_Proinflammatory$GeneName == "APOE"
                               |diff_markers_Proinflammatory$GeneName == "APOC1"
                              |diff_markers_Proinflammatory$GeneName == "C1QA"
                              |diff_markers_Proinflammatory$GeneName == "C1QB"
                               |diff_markers_Proinflammatory$GeneName == "C1QC"
                              |diff_markers_Proinflammatory$GeneName == "FTL" 
                              |diff_markers_Proinflammatory$GeneName == "CCL18"
                              |diff_markers_Proinflammatory$GeneName == "HSPA1B"
                              |diff_markers_Proinflammatory$GeneName == "HSPA6"
                              ,TRUE, FALSE)  
  
#Draw Volcano
q<-ggplot(data=diff_markers_Proinflammatory, aes(x=avg_log2FC, y=-log10(p_val_adj), col=diffexpressed # label= delabel)) +
                                      ))+ 
      geom_point() + 
    theme_minimal()+
  geom_text_repel(aes(avg_log2FC, -log10(p_val_adj)),
     label = ifelse(diff_markers_Proinflammatory$genelabels, diff_markers_Proinflammatory$GeneName, ""),
     box.padding = unit(0.45, "lines"),
     hjust = 1, max.overlaps = Inf) 


q

# Add lines as before...
q2 <- q + geom_vline(xintercept=c(--0, -0), col="black") + xlim(-2, 2)+ 
        geom_hline(yintercept=-log10(0.05), col="black") +  ylim(0, 75) 
q2

# 2. to automate a bit: ceate a named vector: the values are the colors to be used, the names are the categories they will be assigned to:
mycolors <- c("blue", "red", "black")
names(mycolors) <- c("DOWN", "UP", "NO")
q3 <- q2 + scale_colour_manual(values = mycolors)

q3

save_plot(paste0(script_number,prefix,"_Volcano_Proinflammatory_LOvsHi_selec_limit_y"), q3, height = 8,width = 8)



```




















```{r}
sessionInfo()
```
