---
title: "scRNASeq PROGENy"
author: "D. Krauss & O. Fari"
output:
  rmdformats::robobook:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: true
    highlight: tango
  html_document:
    toc: yes
    toc_depth: 4
    df_print: paged
    keep_md: yes
  html_notebook:
    theme: spacelab
    highlight: textmate
    toc: yes
    toc_depth: 4
    df_print: paged
    toc_float: yes
---


***
# Load libraries
```{r setup}
knitr::opts_chunk$set(message = TRUE)
# load required libraries
library(knitr)
library("tidyverse")
library("conflicted")
library("BiocParallel")
library("ggplot2")
library("readr")
library("RColorBrewer")
library("dplyr")
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("count", "dplyr")
conflict_prefer("rowRanges", "MatrixGenerics")
library("ggpubr")
library("tibble")
conflict_prefer("paste", "base")
conflict_prefer("rename", "dplyr")
library("tidyr")
library("org.Hs.eg.db") 
library("org.Mm.eg.db") 
conflict_prefer("desc", "dplyr")

library(Seurat)
library(Matrix)
library(RCurl)
library(scales)
library(cowplot)
library(SingleCellExperiment)
library(AnnotationHub)
library(ensembldb)
library(multtest)
library(glmGamPoi)
library(metap)
library(patchwork)

library(sctransform)
library(SingleR)
library(MAST)
library(DESeq2)
library(Nebulosa)
library(celldex)
library(ggrepel)
library("ComplexHeatmap")
library("circlize")

save_plot <- function(filename, p, width=NULL, height=NULL) {
  if (!is.null(width) && !is.null(height)) {
   # ggsave(file.path(plotsdir, paste0("/",filename, ".png")), plot = p, width = width, height = height)
   # ggsave(file.path(plotsdir, paste0("/",filename, ".svg")), plot = p, width = width, height = height)
    ggsave(file.path(plotsdir, paste0("/",filename, ".pdf")), plot = p, width = width, height = height)
    
  } else {
   # ggsave(file.path(plotsdir, paste0("/",filename, ".png")), plot = p)
  #  ggsave(file.path(plotsdir, paste0("/",filename, ".svg")), plot = p)
    ggsave(file.path(plotsdir, paste0("/",filename, ".pdf")), plot = p)
  }
}
save_RDS <- function(dir, object,filename){
  saveRDS(object, file.path(
    dir,paste0(prefix,"/","_",filename,".RDS")))
}

```




# Load-prepare dataset & add annotation
```{r load data}
load("/srv/workspace/rprojects/USER/Seurat_Myeloid/data/seurat_object_myl.Rdata")

annotation <- data.frame(seurat_clusters = c(0:11),
                         Celltype = c("Macro_C1qc", "Granulo_S100a9",
                                      "Macro_Vegfa", "Granulo_Cxcl2","cDC1_CD209a", "Monos_chil3", 
                                      "DCs_act", "Granulo_Homox1","pDCs", "Mast","Basophilic", "Granulo_Mmp8"))
annotation$seurat_clusters <- as.factor(annotation$seurat_clusters)
metadata <- seurat_object_myl@meta.data %>%
  add_column(Genotype = substr(rownames(seurat_object_myl@meta.data),1,3)) %>%
  add_column(Genotype_short = substr(rownames(seurat_object_myl@meta.data),1,2)) %>%
  add_column(rowIDs = rownames(seurat_object_myl@meta.data))

#metadata$CellGeno <- (paste(metadata$Celltype, "_", metadata$Genotype_short))
  

meta_merge <- left_join(metadata, annotation) %>%
  column_to_rownames("rowIDs") 
meta_merge$CellGeno <- (paste0(meta_merge$Celltype, "_", meta_merge$Genotype_short))




all(seurat_object_myl@meta.data$nCount_RNA == meta_merge$nCount_RNA)
all(rownames(meta_merge) == rownames(seurat_object_myl@meta.data))


seurat_object_myl@meta.data <- meta_merge
# ### workaround color dimplot error: Fehler: Must request at least one colour from a hue palette.
# seurat_object_myl$cell_label <- Idents(seurat_object_myl)
  
  

```

## subset 
```{r subset dataset}
###########################
### subset cluster
#variable <- c("0", "2","5") #here you can put any variable you want to select
#seurat_object_subset <- subset(x = seurat_object_myl, idents = variable)

#### OR
###########################
### subset parameter
# variable2 <- c("wtp", "kop") #"wtp" "kop" "kon" "wtn"
# seurat_object_subset <- subset(seurat_object_myl, subset = Genotype == variable2)




seurat_object_subset <- seurat_object_myl
```





# PROGENy script 
from here https://saezlab.github.io/progeny/articles/ProgenySingleCell.html
R version 4.0.4 (2021-02-15)
```{r PROGENY f. before}
CellsClusters <- data.frame(Cell = names(Idents(seurat_object_subset)), 
                            Celltype = as.character((seurat_object_subset$Celltype)),
                            Genotype = as.character((seurat_object_subset$Genotype_short)),
                            stringsAsFactors = FALSE)
CellsClusters$CellGeno <- (paste(CellsClusters$Celltype, "_", CellsClusters$Genotype))

##Create dataframe for Genotypes
##GenotypeCLusters <- data.frame(Cell = names(Idents(seurat_object_subset)), 
                            ##Genotype = as.character((seurat_object_subset$Genotype)),
                            ##stringsAsFactors = FALSE)

## We compute the Progeny activity scores and add them to our Seurat object
## as a new assay called Progeny. 
data <- progeny(seurat_object_subset, scale=FALSE, organism="Mouse", top=500, perm=1, 
                return_assay = TRUE)

## We can now directly apply Seurat functions in our Progeny scores. 
## For instance, we scale the pathway activity scores. 
data <- Seurat::ScaleData(data, assay = "progeny") 

## We transform Progeny scores into a data frame to better handling the results
progeny_scores_df <- 
  as.data.frame(t(GetAssayData(data, slot = "scale.data", 
                               assay = "progeny"))) %>%
  rownames_to_column("Cell") %>%
  gather(Pathway, Activity, -Cell) 

## We match Progeny scores with the cell clusters.
progeny_scores_df <- inner_join(progeny_scores_df, CellsClusters)

## We summarize the Progeny scores by cellpopulation
summarized_progeny_scores <- progeny_scores_df %>% 
  group_by(Pathway, CellGeno) %>%
  summarise(avg = mean(Activity), std = sd(Activity))


## We prepare the data for the plot
summarized_progeny_scores_df <- summarized_progeny_scores %>%
  dplyr::select(-std) %>%   
  spread(Pathway, avg) %>%
  data.frame(row.names = 1, check.names = FALSE, stringsAsFactors = FALSE) 


paletteLength = 100
myColor = colorRampPalette(c("Darkblue", "white","red"))(paletteLength)
summarized_progeny_scores_df <- summarized_progeny_scores_df %>%
select(c("EGFR","JAK-STAT","PI3K")) 

summarized_progeny_scores_df <- summarized_progeny_scores_df[rev(rownames(summarized_progeny_scores_df)),]

progenyBreaks = c(seq(min(summarized_progeny_scores_df), 0, 
                      length.out=ceiling(paletteLength/2) + 1),
                  seq(max(summarized_progeny_scores_df)/paletteLength, 
                      max(summarized_progeny_scores_df), 
                      length.out=floor(paletteLength/2)))

progeny_hmap = pheatmap::pheatmap(t(summarized_progeny_scores_df#[,-1]
                                    ),
                        fontsize=12, 
                        fontsize_row = 10, 
                        color=myColor,
                        breaks = progenyBreaks,
                        main = "PROGENy (500)",
                        cluster_cols = FALSE,
                        angle_col = 45,
                        treeheight_col = 0, 
                        border_color = NA)

p <- progeny_hmap
save_plot(paste0("EGFR_Pathway_Progeny"), p, height = 10,width = 18)

```




## PROGENy New
https://saezlab.github.io/decoupleR/articles/pw_sc.html
version  R version 4.2.2 (2022-10-31)
```{r PROGENy sc through decoupleR}
library(progeny)
library(decoupleR)
library(pheatmap)
net <- get_progeny(organism = 'mouse', top = 100)
net

# Extract the normalized log-transformed counts
data <- seurat_object_subset
mat <- as.matrix(data@assays$RNA@data)

# Run wmean
acts <- run_wmean(mat=mat, net=net, .source='source', .target='target',
                  .mor='weight', times = 100, minsize = 5)
acts

# Extract norm_wmean and store it in pathwayswmean in data
data[['pathwayswmean']] <- acts %>%
  dplyr::filter(statistic == 'norm_wmean') %>%
  pivot_wider(id_cols = 'source', names_from = 'condition',
              values_from = 'score') %>%
  column_to_rownames('source') %>%
  Seurat::CreateAssayObject(.)

# Change assay
DefaultAssay(object = data) <- "pathwayswmean"

# Scale the data
data <- ScaleData(data)
data@assays$pathwayswmean@data <- data@assays$pathwayswmean@scale.data

##### For summarising
## We can now directly apply Seurat functions in our Progeny scores. 
## For instance, we scale the pathway activity scores. 
#data <- Seurat::ScaleData(data, assay = "pathwayswmean") 

## We transform Progeny scores into a data frame to better handling the results
#progeny_scores_df <- 
  #as.data.frame(t(GetAssayData(data, slot = "scale.data", 
                               #assay = "pathwayswmean"))) %>%
  #rownames_to_column("Cell") %>%
 # gather(Pathway, Activity, -Cell) 

## We match Progeny scores with the cell clusters.
#progeny_scores_df <- inner_join(progeny_scores_df, CellsClusters)

## We summarize the Progeny scores by cellpopulation
#summarized_progeny_scores <- progeny_scores_df %>% 
  #group_by(Pathway, CellGeno) %>%
  #summarise(avg = mean(Activity), std = sd(Activity))


## We prepare the data for the plot
#summarized_progeny_scores_df <- summarized_progeny_scores %>%
  #dplyr::select(-std) %>%   
  #spread(Pathway, avg) %>%
  #data.frame(row.names = 1, check.names = FALSE, stringsAsFactors = FALSE) 

# Change order of WT and KO in plots
data$Genotype_short <- factor(x = data$Genotype_short, levels = c("wt", "ko"))
data$Cell <- factor(x = data$Celltype, levels = c("Macro_C1qc", "Granulo_S100a9",
                                      "Macro_Vegfa", "Granulo_Cxcl2","cDC1_CD209a", "Monos_chil3", 
                                      "DCs_act", "Granulo_Homox1","pDCs", "Mast","Basophilic", "Granulo_Mmp8"))


#####Visualizing
# UMAP Activity Plot-EGFR 
#"lightgrey", "#04103C"

p1 <- DimPlot(data, reduction = "umap", label = TRUE, pt.size = 0.5, split.by = "Genotype_short") + 
  #NoLegend() + 
  ggtitle('Cell types')
p2 <- (FeaturePlot(data, features = c("EGFR"), split.by = "Genotype_short") & #### CHANGE here for plotting a different pathway
  scale_colour_gradient2(low = 'black', mid = 'lightgrey', high = 'red')) +
  ggtitle('EGFR activity')
p1 | p2
p <- p1 + p2
save_plot(paste0("_EGFR_Pathway_Progeny_UMAP"), p, height = 10,width = 25)



# UMAP Activity Plot-PI3K

p1 <- DimPlot(data, reduction = "umap", label = TRUE, pt.size = 0.5, split.by = "Genotype_short") + 
  #NoLegend() + 
  ggtitle('Cell types')
p2 <- (FeaturePlot(data, features = c("PI3K"),pt.size = 0.5, split.by = "Genotype_short") & #### CHANGE here for plotting a different pathway
  scale_colour_gradient2(low = 'black', mid = 'lightgrey', high = 'red')) +
  ggtitle('PI3K activity')
p1 | p2
p <- p1 + p2
save_plot(paste0("_PI3K_Pathway_Progeny_UMAP"), p, height = 10 ,width = 25)
```


##
```{r PROGENy sc through decoupleR Part2}
#subset  cells- UMAP activity
Idents(data) <- data$seurat_clusters

data_subset <- subset(x = data, 
                         idents = c("0","2"))
Idents(data_subset) <- data_subset$Celltype

p1 <- DimPlot(data_subset, reduction = "umap", label = TRUE, pt.size = 0.5, group.by = "Genotype_short") + 
  NoLegend() + ggtitle('Cell types')
p2 <- DimPlot(data_subset, reduction = "umap", label = TRUE, pt.size = 0.5, group.by = "Celltype") + 
  NoLegend() + ggtitle('Cell types')
p3 <- (FeaturePlot(data_subset,  features = c("EGFR"), max.cutoff = 2,  pt.size = 0.5, reduction = "umap", label = T, raster=F) & scale_colour_gradient2(low = '#04103C', mid = 'lightgrey', high = '#e84445')) +
    ggtitle('EGFR activity')
p1 | p2 | p3


p <- p1 | p2 | p3
p

#Plot VlnPlot_scCustom PI3K with stats
install.packages("scCustomize")
library(scCustomize)
Idents(data) <- data$Celltype
colours <- c("grey", "darkred")
p <- VlnPlot_scCustom(seurat_object = data, 
                      features = "PI3K",
                      raster = FALSE,
                      #group.by = "Genotype_short",
                      split.by = "Genotype_short",
                      colors_use = colours,
                      pt.size = 0.01) + stat_summary(fun = median, geom='point', position = position_dodge(0.9), size = 5, 
                                                     colour = "dark blue") + stat_compare_means(label = "p.format")
p

save_plot(paste0("_violin_PI3K_activity_Progeny"), p, height = 10,width = 16)

#smooth violin Pi3k
q <- VlnPlot(object = data,
             features = "PI3K",
             split.by = "Genotype_short",
             cols = colours, pt.size = 0) + stat_summary(
               fun = median, geom='point', 
               position = position_dodge(0.9), size = 5, 
                                              colour = "light blue") + stat_compare_means(label = "p.signif")

q

save_plot(paste0("_violin_PI3K_activity_Progeny_smooth"), q, height = 10,width = 16)


#Plot VlnPlot_scCustom JAK-STAT with stats
library(scCustomize)
library(ggpubr)
Idents(data) <- data$seurat_clusters
colours <- c("grey", "darkred")
p <- VlnPlot_scCustom(seurat_object = data, 
                      features = "JAK-STAT",
                      raster = FALSE,
                      #group.by = "Genotype_short",
                      split.by = "Genotype_short",
                      colors_use = colours,
                      pt.size = 0.01) + stat_summary(fun = median, geom='point', position = position_dodge(0.9), size = 5, 
                                                     colour = "dark blue") + stat_compare_means(label = "p.signif")
p

save_plot(paste0("_violin_JAK-STAT_activity_Progeny"), p, height = 10,width = 16)

#smooth violin jak
q <- VlnPlot(object = data,
             features = "JAK-STAT",
             split.by = "Genotype_short",
             cols = colours, pt.size = 0) + stat_summary(
               fun = median, geom='point', 
               position = position_dodge(0.9), size = 5, 
                                              colour = "light blue") + stat_compare_means(label = "p.signif")
q


save_plot(paste0("_violin_JAK-STAT_activity_Progeny_smooth"), q, height = 10,width = 16)

#Plot VlnPlot_scCustom EGFR
library(scCustomize)
Idents(data) <- data$seurat_clusters
p <- VlnPlot_scCustom(seurat_object = data, 
                      features = "EGFR",
                      raster = FALSE,
                      #group.by = "Genotype_short",
                      split.by = "Genotype_short",
                       colors_use = colours,
                      pt.size = 0.01) + stat_summary(fun = median, geom='point', position = position_dodge(0.9), size = 5, 
                                                     colour = "dark blue") + stat_compare_means(label = "p.signif")
p
save_plot(paste0("_violin_EGFR_activity_Progeny"), p, height = 10,width = 16)

#smooth violin egfr
q <- VlnPlot(object = data,
             features = "EGFR",
             split.by = "Genotype_short",
             cols = colours, pt.size = 0) + stat_summary(
               fun = median, geom='point', 
               position = position_dodge(0.9), size = 5, 
                                              colour = "light blue") + stat_compare_means(label = "p.signif")

q

save_plot(paste0("_violin_EGFR_activity_Progeny_smooth"), q, height = 10,width = 16)

Idents(data) <- data$CellGeno
# Extract activities from object as a long dataframe
df <- t(as.matrix(data@assays$pathwayswmean@data)) %>%
  as.data.frame() %>%
  mutate(cluster = Idents(data)) %>%
  pivot_longer(cols = -cluster, names_to = "source", values_to = "score") %>%
  group_by(cluster, source) %>%
  summarise(mean = mean(score))


# Transform to wide matrix
top_acts_mat <- df %>%
  pivot_wider(id_cols = 'cluster', names_from = 'source',
              values_from = 'mean') 

top_acts_mat <- top_acts_mat %>%
    add_column(ordering = as.character(top_acts_mat$cluster))

Selected_Cluster <- c("Macro_C1qc_wt","Macro_C1qc_ko", "Macro_Vegfa_wt", "Macro_Vegfa_ko")# 


top_acts_mat <- top_acts_mat %>%
  arrange(desc(ordering), ) %>%
  select(c("EGFR","JAK-STAT","PI3K")) %>%
  filter(cluster %in% Selected_Cluster) %>% 
  column_to_rownames('cluster') %>%
  as.matrix()


# Choose color palette
palette_length = 100
my_color = colorRampPalette(c("Darkblue", "white","red"))(palette_length)

my_breaks <- c(seq(-2, 0, length.out=ceiling(palette_length/2) + 1),
               seq(0.05, 2, length.out=floor(palette_length/2)))



# Plot Heatmap
p <- pheatmap::pheatmap(top_acts_mat, border_color = NA, color=my_color, breaks = my_breaks, cluster_rows = FALSE) 

p
save_plot("_pheatmap_macros", p, height = 10, width = 18)





## Human??
### PLOT WATERFALL
f_contrast_acts <- acts %>%
  filter(statistic == 'norm_wmean')

# Plot
ggplot(f_contrast_acts, aes(x = reorder(source, score), y = score)) + 
    geom_bar(aes(fill = score), stat = "identity") +
    scale_fill_gradient2(low = "darkblue", high = "indianred", 
        mid = "whitesmoke", midpoint = 0) + 
    theme_minimal() +
    theme(axis.title = element_text(face = "bold", size = 12),
        axis.text.x = 
            element_text(angle = 45, hjust = 1, size =10, face= "bold"),
        axis.text.y = element_text(size =10, face= "bold"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
    xlab("Pathways")








# Extract activities from object as a long dataframe
df <- t(as.matrix(data@assays$pathwayswmean@data)) %>%
  as.data.frame()
hist(df$EGFR, breaks = 1000)

# Extract activities from object as a long dataframe for myeloid cells
df <- t(as.matrix(data@assays$pathwayswmean@data)) %>%
  as.data.frame()
hist(df$EGFR, breaks = 1000)

ordered_egfr = df$EGFR[order(df$EGFR,decreasing = F)]
plot(x = 1:length(ordered_egfr), y = ordered_egfr)

length(df$EGFR)
#table(df$EGFR)

length(which(df$EGFR >= 0))
df_ordered <- df %>% arrange(EGFR)
  
quantile(x = df_ordered$EGFR, type = 4)



egfr.cluster <- df %>%
  add_column(EGFR_positive = df$EGFR >= 0.59) %>%
  add_column(EGFR_negative = df$EGFR <= -0.71) %>%
  mutate(EGFR_pathway = case_when(
    EGFR_positive == TRUE ~ "up",
    EGFR_negative == TRUE ~ "down",
    EGFR_positive == FALSE ~ "else"))
all(rownames(egfr.cluster) == rownames(data@meta.data))

data$EGFR_pathway_activity <- egfr.cluster$EGFR_pathway


############## plot again after classification
DimPlot(data, 
        pt.size = 0.5,
        reduction = "umap", 
        group.by = "EGFR_pathway_activity",
        label = F,
        cols = c("#d9d9d9","lightgrey","#ff0000" )
        ) &
    ggtitle('EGFR activity')




p1 <- DimPlot(data, 
              reduction = "umap", 
              label = TRUE, pt.size = 0.5, 
              group.by = "Celltype") + 
  NoLegend() + ggtitle('Cell types')
p2 <- DimPlot(data, 
              reduction = "umap", 
              label = TRUE, 
              pt.size = 0.5, 
              group.by = "Genotype") + 
  NoLegend() + ggtitle('Cell types')
p3 <- DimPlot(data, 
              pt.size = 0.5,
              reduction = "umap", 
              group.by = "EGFR_pathway_activity",
              label = F,
              cols = c("#d9d9d9","lightgrey","#ff0000")) &
    ggtitle('EGFR activity')
p1 | p2 | p3

# p <- p1 | p2 | p3
# save_plot(paste0(script_number,prefix,"_EGFR_Pathway_Progeny_Myeloid_classification"), p, height = 10,width = 18)

prop.table(table(data$EGFR_pathway_activity,data$Celltype ), margin = 2)
prop.table(table(data$EGFR_pathway_activity,data$Genotype ), margin = 2)
```







###PROGENy statistics - IROS suggestion -
https://bioconductor.org/packages/release/bioc/vignettes/decoupleR/inst/doc/pw_bk.html
```{r}
selected_clust <- subset(data, idents = 0 )


mat <- as.matrix(selected_clust@assays$RNA@data)

design <- tibble(sample = rownames(selected_clust@meta.data),
                 condition = selected_clust@meta.data$Genotype)
all(colnames(mat)==design$sample)
# # Extract t-values per gene
# deg <- data$limma_ttop %>%
#     select(ID, t) %>% 
#     filter(!is.na(t)) %>% 
#     column_to_rownames(var = "ID") %>%
#     as.matrix()
# head(deg)


net <- get_progeny(organism = 'mouse', top = 100)
net
# Run wmean
sample_acts <- run_wmean(mat=mat, net=net, .source='source', .target='target',
                  .mor='weight', times = 100, minsize = 5)
sample_acts



# # Extract t-values per gene
# deg <- data$limma_ttop %>%
#     select(ID, t) %>% 
#     filter(!is.na(t)) %>% 
#     column_to_rownames(var = "ID") %>%
#     as.matrix()
# head(deg)
```



### Progeny f. RNASeq for statistics

### PROGENy
```{r PROGENy all, error=T}
# Extract the normalized log-transformed counts
data <- seurat_object_subset
gene_expr <- as.matrix(data@assays$RNA@data)









# rld <- rlog(dds, blind=T)
# gene_expr <- assay(rld)
# gene_expr <- gene_expr[,which(colnames(gene_expr)%in% colnames(dds))]

library(progeny)
pathways = progeny(gene_expr, 
                   scale=FALSE,
                   organism = "Mouse")

rownames(pathways)
rownames(pathways) <- seurat_object_subset$Genotype_short

controls =  rownames(pathways) == unique(seurat_object_subset$Genotype_short)[1]
ctl_mean = apply(pathways[controls,], 2, mean)
ctl_sd = apply(pathways[controls,], 2, sd)
pathways_stat = t(apply(pathways, 1, function(x) x - ctl_mean))
pathways_stat = apply(pathways, 1, function(x) x / ctl_sd)

result = apply(pathways_stat, 1, function(x) {
    broom::tidy(lm(x ~ !controls)) %>%
        filter(term == "!controlsTRUE") %>%
        dplyr::select(-term)
})
res.table <- mutate(bind_rows(result), pathway=names(result))
```








## Seurat Myeloid Object- Gene Expression
```{r sc-myeloid clusters- gene expression}
# Visualize QC metrics as a violin plot
library(Seurat)
library(SeuratData)
library(cowplot)
library(patchwork)

# Look for RNA counts and mitochondrial features
VlnPlot(data, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)


# Identify conserved cell type markers
DefaultAssay(data) <- "RNA"
nk.markers <- FindConservedMarkers(data, ident.1 = 0, grouping.var = "Genotype_short", verbose = FALSE)
head(nk.markers)

# Identify differentially expressed genes- between genotypes
# Plot Dot plot 

markers.to.plot <- c("Il1a", "Il1b", "Ccr1", "S100a9", "Cd86", "Cxcl1", "Lyz2")

Idents(data) <- "CellGeno"
levels(Idents(data)) <- rev(sort(paste(levels(Idents(data)))))
DotPlot(data, features = rev(markers.to.plot), dot.scale = 4, cols = c("red"#, "white
)#,split.by = "Genotype_short"
) + RotatedAxis()

Idents(data) <- data$seurat_clusters





#Identify differential expressed genes- between genotypes
#Plot Volcano Plot customize- for genes of interest

VlnPlot_scCustom(data, features = c("Fos", "Jun", "Myc", "Sta3", "Akt3", "Nfkb1", "Nfkb2"), 
                 idents =  c("0", "1", "2", "3", "4", "5"),
                 split.by = "Genotype_short") + stat_summary(fun = median, geom='point', position = position_dodge(0.9), size = 3, colour = "magenta") + stat_compare_means(label = "p.signif")

#other genes "Il1a", "Il1b", "Ccr1", "S100a9", "Lyz2", "Cd86", "Cd80", "Myd88", "Nfkbia"

# Identify EGFR downstream targets
# Plot Volcano Plot- for downstream EGFR genes
q <- VlnPlot(data, features = c("Ccr1", "Ccr2", "Ccl3", "Ccl4", "Ccl5", "Cxcl10", "Ccl9", "Cx3cr1", "Hmgb1", "Ctsg", "App", "Lmnb1"), idents =  c("0", "1", "2", "3", "4", "5", "6", "7"), split.by = "Genotype_short")
q
save_plot("_Monos_Chemotaxis", q , height = 10, width = 18)

#ERBB1 "Src", "Akt3", "Pik3r5", "Mtor", "Egfr", "Jun", "Myc", "Prkca", "Sos1"
# MM6218 "Ifi206", "Cdk6", "Csf1", "Fasn", "Hoxa7", "Jun", "Myc", "Vegfa", "Ifi204", "Ifi207", "Pparg", "Gpr68"
# MM15900 "Atp5b", "mt-Atp6", "mt-Nd1", "mt-Nd4", "Ndufa2", "Ndufa9", "Dmac2l"
# MM1516 "Tnfrsf1a", "Tradd", "Casp8", "Casp2", "Casp2", "Ripk1", "Casp3", "Fadd", "Lmnb1"
# MM4303 "Ccr1", "Ccr2", "Ccl3", "Ccl4", "Ccl5", "Cxcl10", "Ccl9", "Cx3cr1", "Hmgb1", "Ctsg", "App"

# Plot Feature plot for individual genes 
FeaturePlot(data, features = c("Il1b"), dims = c(1, 2), pt.size = 0.5, split.by = "Genotype_short")


```

























## PSEUDOBULK
as described in https://hbctraining.github.io/scRNA-seq/lessons/pseudobulk_DESeq2_scrnaseq.html


Rational
- differential expression analysis
- functions of Seurat available, however p-values often inflated (each cell treated as sample)

Steps:
1. Subsetting cells f. cell types of interest
2. Extracting raw counts after QC
3. Aggregating counts & metadata
4. Perform DE analysis

New literature
Benchmarking methods for detecting differential statesbetween conditions from multi-subject single-cellRNA-seq data  Junttila, Smolander, Elo
https://doi.org/10.1093/bib/bbac286

Confronting false discoveries in single-cell differential expression

```{r Pseudobulk DE all cells analysis, eval=FALSE, include=FALSE}
library(scater)
library(edgeR)
library(Matrix.utils)
library(magrittr)
library(purrr)
library(reshape2)
library(pheatmap)
library(apeglm)
library(png)

seurat_object_myl$AnimalID <- substr(rownames(seurat_object_myl@meta.data),4,7)
seurat_obj_normalized <- seurat_object_myl



counts <- seurat_obj_normalized@assays$RNA@counts
metadata <- seurat_obj_normalized@meta.data

metadata <- metadata %>% #as.factor(sce$condition_id) 
  mutate(AnimalID = case_when(
    AnimalID == 3428 ~ "A",
    AnimalID == 3429 ~ "A",
    AnimalID == 3430 ~ "C",
    AnimalID == 3431 ~ "C")) %>%
  mutate(Celltype = gsub('_','-',metadata$Celltype))



table(metadata$Genotype_short)
table(metadata$AnimalID)
table(metadata$Celltype)


#metadata$cluster_id <- factor(seurat_obj_normalized@active.ident)
#or
metadata$cluster_id <- factor(metadata$Celltype)


sce <- SingleCellExperiment(assays = list(counts = counts), colData = metadata)


sce$sample_id <- as.factor(sce$AnimalID) 
sce$group_id <- as.factor(sce$Genotype_short) #as.factor(sce$condition)
groups <- colData(sce)[,c("cluster_id","sample_id")]

### define groups to compare
#sce <- sce[,c(which(sce$group_id == "ko"),which(sce$group_id == "wt"))]


# metrics for aggregation
kids <- purrr::set_names(levels(sce$cluster_id))
kids 
nk <- length(kids)
nk
sids <- purrr::set_names(levels(sce$sample_id))
sids
ns <- length(sids)
ns
# Generate sample level metadata
## Determine the number of cells per sample
table(sce$sample_id)

## Turn named vector into a numeric vector of number of cells per sample
n_cells <- as.numeric(table(sce$sample_id))

## Determine how to reoder the samples (rows) of the metadata to match the order of sample names in sids vector
m <- base::match(sids, sce$sample_id)

## Create the sample level metadata by combining the reordered metadata with the number of cells corresponding to each sample.
ei <- data.frame(colData(sce)[m, ], 
                  n_cells, row.names = NULL) %>% 
                select(-"cluster_id")
ei

# remove poor quality cells /if not already happened
dim(sce)
QC <- perCellQCMetrics(sce)
sce$is_outlier <- isOutlier(
  metric = QC$total,
  nmads = 2, type = "both", log = T)

sce <- sce[ ,!sce$is_outlier]
dim(sce)

sce <- sce[rowSums(counts(sce) > 1) >=10,]
dim(sce)

# count aggregation to sample level

groups <- colData(sce)[, c(
  "cluster_id", 
  "sample_id")]
pb <- aggregate.Matrix(t(counts(sce)),
                       groupings = groups, fun = "sum")

class(pb)
dim(pb)
pb[1:6,1:6]

#per cell-type analysis
# Not every cluster is present in all samples; create a vector that represents how to split samples
splitf <- sapply(stringr::str_split(rownames(pb),
                                    pattern = "_",
                                    n = 2),
                 `[`, 1)
# Turn into a list and split the list into components for each cluster and transform, so rows are genes and columns are samples and make rownames as the sample IDs
pb <- split.data.frame(pb,
                       factor(splitf)) %>%
        lapply(function(u)
                set_colnames(t(u),
                             stringr::str_extract(rownames(u), "(?<=_)[:alnum:]+")))

class(pb)
str(pb)
options(width = 100)
table(sce$cluster_id, sce$sample_id)




# prep. data.frame for plotting
get_sample_ids <- function(x){
        pb[[x]] %>%
                colnames()
}

de_samples <- purrr::map(1:length(kids), get_sample_ids) %>%
        unlist()

# Get cluster IDs for each of the samples

samples_list <- map(1:length(kids), get_sample_ids)

get_cluster_ids <- function(x){
        rep(names(pb)[x],
            each = length(samples_list[[x]]))
}

de_cluster_ids <- map(1:length(kids), get_cluster_ids) %>%
        unlist()


# Create a data frame with the sample IDs, cluster IDs and condition

gg_df <- data.frame(cluster_id = de_cluster_ids,
                    sample_id = de_samples)

gg_df <- left_join(gg_df, ei[, c("sample_id", "group_id")]) 


metadata <- gg_df %>%
        dplyr::select(cluster_id, sample_id, group_id) 
metadata$cluster_id <- as.factor(metadata$cluster_id)
        
metadata 

# Generate vector of cluster IDs
clusters <- levels(metadata$cluster_id)
clusters
# 
# Subset the metadata to only the xxx
cluster_metadata <- metadata[which(metadata$cluster_id == clusters[8]), ]


head(cluster_metadata)

# Assign the rownames of the metadata to be the sample IDs
rownames(cluster_metadata) <- cluster_metadata$sample_id
head(cluster_metadata)

# Subset the counts to only the XX
counts <- pb[[clusters[8]]]

cluster_counts <- data.frame(counts[, which(colnames(counts) %in% rownames(cluster_metadata))])

# Check that all of the row names of the metadata are the same and in the same order as the column names of the counts in order to use as input to DESeq2
all(rownames(cluster_metadata) == colnames(cluster_counts))

```

```{r dds on single comparison all cells, eval=FALSE, include=FALSE}

### create DESeq2 Object
design_formula <- as.formula(paste0("~ ", 
                                    "group_id"))

dds <- DESeqDataSetFromMatrix(cluster_counts,
                              colData = cluster_metadata,
                              design = design_formula)

# Transform counts for data visualization
rld <- rlog(dds, blind=TRUE)

# Plot PCA

DESeq2::plotPCA(rld, intgroup = "group_id")

# Extract the rlog matrix from the object and compute pairwise correlation values
rld_mat <- assay(rld)
rld_cor <- cor(rld_mat)

# Plot heatmap
q <- pheatmap::pheatmap(rld_cor, annotation = cluster_metadata[, c("group_id"), drop=F])
q

### PROGENy
#rld <- rlog(dds, blind=T)
gene_expr <- assay(rld)
gene_expr <- gene_expr[,which(colnames(gene_expr)%in% colnames(dds))]

library(progeny)
pathways = progeny(gene_expr, 
                   scale=FALSE,
                   organism = "Mouse")

rownames(pathways)
rownames(pathways) <- dds$group_id

controls =  rownames(pathways) == dds$group_id[1]
ctl_mean = apply(pathways[controls,], 2, mean)
ctl_sd = apply(pathways[controls,], 2, sd)
pathways_stat = t(apply(pathways, 1, function(x) x - controls))
pathways_stat = apply(pathways, 1, function(x) x / controls)

result = apply(pathways_stat, 1, function(x) {
    broom::tidy(lm(x ~ !controls)) %>%
        filter(term == "!controlsTRUE") %>% 
        dplyr::select(-term)}) # gives error because of Inf values

res.table <- mutate(bind_rows(result), pathway=names(result))

plot.pathways <- progeny(gene_expr,
                    scale = TRUE, 
                    organism = "Mouse", 
                    top = 100,
                    perm = 1, 
                    verbose = FALSE)

## select significant pathways
res.sig <- res.table %>%
  filter(p.value < 0.1)
selected <- res.sig$pathway


myColor = colorRampPalette(brewer.pal(11, "BrBG"))(100) # green from here
a <- myColor[51:100]
myColor = colorRampPalette(brewer.pal(11, "RdBu"))(100) # get the red from here
b <- myColor[1:50]
myColor <- c(b,a)


plot.pathways2 <- plot.pathways[,selected]
p <- pheatmap::pheatmap(plot.pathways,
         fontsize=14, 
         show_rownames = T,
         cluster_rows = T,
         cluster_cols = T,
         show_colnames = T,
         #fontsize_row = 10 - nrow(plot.pathways) / 15,
         legend = T,
         color= rev(myColor), 
         main = "PROGENy", 
         angle_col = 45, 
         #treeheight_col = 0,  
         border_color = NA)
p
#save_plot(paste0(script_number,prefix,"PROGENy_all"), p)
```




